---
title: "1. Identify Dynamic Genes and Generate Master Table."
output: html_document
---

#### 1. Load TrendCatcher R package.

```{r, results='hide', message=FALSE}
library(TrendCatcher)
```

#### 2. Now read the demo count table.

TrendCatcher **requires** the input count table in CSV file and with the column name in format as "ProjectName_Time_Rep1". 
And the first column as gene symbol or gene ensembl. 

```{r}
example.file.path<-system.file("extdata", "Brain_DemoCountTable.csv", package = "TrendCatcher")
tb<-read.csv(example.file.path, row.names = 1)
head(tb)
```


#### 2. Now read the demo count table and run TrendCatcher. 

This function will take few minutes to finish running. 

```{r, results='hide', eval=FALSE}
example.file.path<-system.file("extdata", "Brain_DemoCountTable.csv", package = "TrendCatcher")

master.list<-run_TrendCatcher(count.table.path = example.file.path,
baseline.t = 0,
time.unit = "h",
min.low.count = 1,
para.core.n = NA,
dyn.p.thres = 0.05)
```

#### 3. Check master.list in detail.

To save time of running, we already put the output master.list object in the '/inst/extdata' folder. You can simply load it into your environment.

First check what elements are included in the master.list list object. 

- "time.unit" is the study time unit, for example "h" represents hour. 
- "baseline.t" is the baseline time. Here it is hour 0.
- "t.arr" is the time array in temporal study. Here is 0, 6, 24, 48, 72, 168h.
- "Project.name" is the project name. Here is "B".
- "raw.df" is the count table ordered by time and replicate ID. 
- "fitted.count" is the ANOVA smooth model fitted count number from replicates.

```{r}
demo.master.list.path<-system.file("extdata", "BrainMasterList.rda", package = "TrendCatcher")
load(demo.master.list.path)
names(master.list)
```

```{r}
print(c(master.list$time.unit, master.list$baseline.t))
```

```{r}
master.list$t.arr
```
```{r}
master.list$Project.name
```

```{r}
head(master.list$raw.df)
```

The fitted.count table each column represents.

- Gene, gene ensembl or gene symbol
- Time, time measured.
- Fit.Count, the ANOVA smooth model fitted count number.
- mu, the estimated mean count.
- disp, dispersion estimated.
- t.p.val, the p-value at each time point.
- dyn.p.val, the combined dynamic p-value.
- dyn.p.val.adj, the adjusted p-value from all hypothesis testing. 

```{r}
head(master.list$fitted.count)
```
For the master table, each column represents.

- Gene, gene ensembl or gene symbol.
- pattern, the gene trajectory master-pattern type.
- start.idx, the index of where the corresponding trend start in t.arr.
- end.idx, the index of where the corresponding trend end in the t.arr.
- dynTime, the time where t.p.val <=0.05.
- dynSign, "-" means down, "+" means up.
- start.t, the time of where the corresponding trend start.
- end.t, the time of where the corresponding trend end.
- pattern_str, the string ready for print indicates the gene trajectory sub-pattern type.


```{r}
head(master.list$master.table)
```



















